---
title: "p8105_hw6_ys3766"
author: "Yifan Shi"
date: "2024-12-03"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Question 1
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Generate 5000 bootstrap samples
```{r}
set.seed(123)
bootstrap_samples <- modelr::bootstrap(weather_df, 5000)
```

Fit models and compute metrics for each bootstrap sample
```{r}
bootstrap_analysis <- bootstrap_samples %>%
  mutate(
    linear_model = map(strap, ~ lm(tmax ~ tmin, data = as_tibble(.x))),
    r_squared_values = map_dbl(linear_model, function(model) glance(model)[["r.squared"]]),
    log_coefficients_product = map_dbl(linear_model, function(model) {
      coefficients <- tidy(model) %>% pull(estimate)
      log(coefficients[1] * coefficients[2])
    })
  ) %>%
  select(r_squared_values, log_coefficients_product)
```

Plot distributions of r-squared and log(beta0 * beta1)
```{r}
bootstrap_analysis %>%
  pivot_longer(cols = everything(), names_to = "metric", values_to = "value") %>%
  ggplot(aes(x = value, fill = metric)) +
  geom_density(alpha = 0.5) +
  facet_wrap(vars(metric), scales = "free", ncol = 1) +
  labs(
    title = "Bootstrap Distributions of R² and log(β0 × β1)",
    x = NULL,
    y = "Density"
  )
```

The upper plot of log(beta0 x beta1) is symmetrically distributed, centered around 2.00, with values ranging from 1.95 to 2.05. This shows moderate variability in the regression coefficients across samples, but supports the stability of the model parameters and a good over all fit. 

The lower plot of R^2 shows a symmetric distribution concentrating between 0.88 and 0.93, with a peak at around 0.91. This indicates a consistently strong model fit across the bootstrap samples, with low variability, suggesting the model reliably explains the variance in tmax using tmin. 

95% CI
```{r}
bootstrap_analysis %>%
  summarise(
    r_squared_lower = quantile(r_squared_values, 0.025),
    r_squared_upper = quantile(r_squared_values, 0.975),
    log_coefficients_product_lower = quantile(log_coefficients_product, 0.025),
    log_coefficients_product_upper = quantile(log_coefficients_product, 0.975)
  ) %>% knitr::kable(digits = 3)
```


